FISTA with adaptive discretisations

### 2021-01-11 paper.tex
Update paper
Some corrections given by Antonin and rewriting of the proof of Lemma A.4

\paragraph{Step 5: Sufficiently slow refinement}
\begin{adaptiveStep}\label{step: step 5 sufficiently slow}
	The result of \Cref{step: step 4 sufficiently fast} is sufficient to prove convergence, but not yet a rate. If the subspaces refine too quickly, then the influence of $\norm{\varf0^*}=\infty$ will slow the rate of convergence. If the resolution is too coarse then we are overfitting to the discrete minimiser, but if the resolution is too fine then convergence of the energy is slow. \Cref{thm: sufficiently slow} balances these two factors in an optimal way for \cref{alg: refining FISTA} resulting in a convergence rate of 
	$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} \lesssim \frac{N^{2\epsilon}}{N^2}$$
	for all $N\in\F N$, some $\epsilon\in[0,1)$ depending on $\vars3_{\Varf0},\vars3_{\Func0}$. In particular, if $\varf0^*\in\F H$ then we recover the classical rate with $\epsilon=0$.
\end{adaptiveStep}

@@

	Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\spcf0^n$ should be refined as soon as the value $\Func0(\varf0_n)$ is sufficiently close to $\Func0(\varf2_n)$. This is analysed in \Cref{thm: stronger exponential FISTA convergence} which shows convergence results of the form
	$$ \min_{n\leq N}\Func0_0(\varf0_n) \lesssim \frac{N^{2\epsilon}}{N^2}$$
	for all $N\in\F N$, with the same $\epsilon\in[0,1)$ from \Cref{step: step 5 sufficiently slow}. The penalty for accelerating the refinement time is a potential loss of stability in $\Func0(\varf0_n)$, however, the asymptotic rate is equivalent and this behaviour has not been seen in numerical experiments.
\end{adaptiveStep}

@@

\begin{lemma}\label{thm: mini exponential FISTA convergence}
	\Copy{thm: mini exponential FISTA convergence}{
		Let $\varf0_n$, $\varf1_n$ be generated by \Cref{alg: refining FISTA}, $(n_k\in\F N)_{k=0}^\infty$ be a monotone increasing sequence, and define
		$$\tilde{\spcf0}^k \coloneqq \spcf0^{n_k}, \qquad \tilde{\varf2}_k \in \argmin_{\varf0\in\tilde{\spcf0}^k} \Func0(\varf0).$$
		If
		$$ \tilde{\varf2}_k \in \spcf0^n \qquad\text{ for all }\qquad n_k\leq n<n_{k+1},\ k\in\F N,$$
		then for all $K\in\F N$, $n_K\leq N<n_{K+1}$ we have
		\begin{multline*}
			\vart1_N^2\Func0_0(\varf0_N) + \sum_{n=1}^{N-1}\rho_n \Func0_0(\varf0_n) + \frac{\norm{\varf1_N-\tilde{\varf2}_K}^2}{2} \leq C + \frac{\norm{\tilde{\varf2}_K}^2}{2} + \frac{(N+1)^2-n_K^2}{2}\Func0_0(\tilde{\varf2}_K) \\+ \sum_{k=1}^K \frac{n_k^2-n_{k-1}^2}{2}\Func0_0(\tilde{\varf2}_{k-1}) + \IP{\varf1_{n_k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}
		\end{multline*}
		where $C = \frac{\norm{\varf0_0-\tilde{\varf2}_0}^2-\norm{\tilde{\varf2}_0}^2}{2}$.
	}
\end{lemma}
\noindent The proof is given in \cref{app:thm: mini exponential FISTA convergence}. The introduction of $n_k$ has greatly compressed the expression of \Cref{thm: mini FISTA convergence}. On the right-hand side, we now only consider $\Func0_0$ evaluated on the sequence $\tilde{\varf2}_k$ and there are $K$ elements to the sum rather than $N$.

@@

\begin{lemma}\label{thm: sufficiently fast}
	\Copy{thm: sufficiently fast}{
		Suppose $\spcf0^n, \varf0_n, \varf1_n$ and $n_k$ satisfy the conditions of \Cref{thm: mini exponential FISTA convergence} and $\{\tilde{\spcf0}^k\}$ forms an $(\vars3_{\Varf0},\vars3_{\Func0})$-discretisation for $\Func0$.
		If either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty,$
		\end{itemize}
		then
		$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2}$$
		for all $n_K\leq N< n_{K+1}$.}
\end{lemma}

@@

\begin{lemma}\label{thm: sufficiently slow}
	\Copy{thm: sufficiently slow}{
		Suppose $\varf0_n$ and $n_k$ are sequences satisfying 
		$$\Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} \qquad\text{where} \qquad  n_K^2\gtrsim \vars3_{\Func0}^K\vars3_{\Varf0}^{2K},$$
		then 
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\epsilon)}}\qquad \text{ where } \qquad \epsilon = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}.$$}
\end{lemma}
\noindent The proof is given in \cref{app:thm: sufficiently slow}.

@@

		\bigbreak\noindent Suppose that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\simeq \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$,
		\end{itemize}
		then
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\epsilon)}}\qquad \text{ where } \qquad \epsilon = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}$$
		uniformly for $N\in\F N$.
	}
\end{theorem}
\noindent This theorem is very easy to implement and requires very little knowledge of how to estimate $\Func0_0(\varf0_n)$. So long as $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be computed analytically, choosing $\tilde{\spcf0}^k$ to be uniform discretisations and $\tilde{\spcf0}^k = \spcf0^{n_k}=\ldots= \spcf0^{n_{k+1}-1}$ will give the stated convergence rate.

@@

\Cref{thm: sufficiently fast} gives a sufficient condition on $n_k$ for converging at the rate $O(N^{2(\epsilon-1)})$ but it is not necessary. Indeed for $n\leq n_k$ the result shows 
$$\Func0_0(\varf0_n) \geq \min_{\varf0\in\spcf0^n}\Func0_0(\varf0) = O(\vars3_{\Func0}^{-k}) = O(n^{2(\epsilon-1)})$$
which suggests that to converge faster than $n^{2(\epsilon-1)}$ requires choosing smaller $n_k$. As an example, in \Cref{sec: wavelet examples} we will see \cref{alg: refining FISTA} can converge at a linear rate, although this is not possible without adaptive refinement times. On the other hand, choice of spatial adaptivity has no impact on rate but can impact computational efficiency. It will be permitted to use greedy discretisation techniques so long as each collection $\{\spcf0^n \st n_k\leq n<n_{k+1}\}$ satisfies a monotonicity condition.

@@

		\bigbreak\noindent Suppose there exist $\vars3_{\Varf0},\vars3_{\Func0}\geq1$ such that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$
		\end{itemize}
		and both
		$$ \norm{\tilde{\varf2}_k}\lesssim \vars3_{\Varf0}^k \qquad\text{and}\qquad \Func0_0(\varf0_{n_K-1}) \lesssim \vars3_{\Func0}^{-K}.$$
		Whenever these conditions on $n_k$ are satisfied, then
		$$\min_{n\leq N}\Func0_0(\varf0_n) \lesssim \frac{1}{N^{2(1-\epsilon)}}\qquad \text{ where } \qquad \epsilon = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}$$
		uniformly for $N\in\F N$.
	}
\end{theorem}
\noindent The proof is given in \cref{app:thm: stronger exponential FISTA convergence}. If we directly compare \Cref{thm: exponential FISTA convergence,thm: stronger exponential FISTA convergence}, we note that the convergence rate is the same in both theorems but the price for better adaptivity is a slightly weaker stability guarantee (i.e. the addition of the min on the left-hand side). In \Cref{thm: exponential FISTA convergence}, as in the original FISTA algorithm, the sequence $\Func0(\varf0_n)$ is not monotone but the magnitude of oscillation is guaranteed to decay in time. This behaviour is boundedness is lost in \cref{thm: stronger exponential FISTA convergence} although it still holds on a subsequence. Although we do not prove it here, it can be shown that the stronger condition 
\begin{equation}
	\tilde{\varf2}_k \in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k+1} \cap\ldots\cap \spcf0^{n_{k+1}-1}\cap\ldots\cap \spcf0^N
\end{equation}
is sufficient to restore the stronger guarantee on $\Func0_0(\varf0_n)$. Again, monotonicity of $\spcf0^n$ corresponds with improved stability of \cref{alg: refining FISTA}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage {\Huge\todo{I don't know if this section should remain}}
\section{General examples}\label{sec: general examples}
In this work we consider the Lasso problem to be both a motivating example and source of numerical examples, however, \Cref{alg: refining FISTA} is much more broadly applicable. The aim of this section is to justify this claim by demonstrating that the constants $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be easily computed in many examples.

@@

\newpage 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Inserting the computed rates into \Cref{thm: exponential FISTA convergence} or \Cref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
\begin{equation}\label{eq: Lasso energy rate}
	\epsilon = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2} = \frac{d}{1+d} \quad \implies \quad \Func0_0(\varf0_n) \lesssim n^{-2(1-\epsilon)} = n^{-\frac{2}{1+d}}.
\end{equation}
This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\sim (\vars3_{\Func0}\vars3_{\Varf0}^2)^k$ we expect the resolution to be
\begin{equation}\label{eq: Lasso resolution rate}
	2^{-k} = \left(\vars3_{\Func0}\vars3_{\Varf0}^2\right)^{\frac{k}{1+d}} \sim n^{-\frac{2}{1+d}}.
\end{equation}

@@

\begin{lemma}[\Cref{thm: mini exponential FISTA convergence}]\label{app:thm: mini exponential FISTA convergence}
	\Paste{thm: mini exponential FISTA convergence}
\end{lemma}
\begin{proof}
	This is just a telescoping of the right-hand side of \eqref{app:eq: FISTA inequality} with the introduction of $n_k$ and simplification $\varf2_n = \tilde{\varf2}_k$,
	\begin{multline}
		\tfrac12\norm{\varf2_N}^2 + \sum^N_{n=1} \vart1_n\Func0_0(\Pi_n\varf2_n) + \IP{\varf1_{n-1}}{\varf2_{n-1}-\varf2_n} = \tfrac12\norm{\tilde{\varf2}_K}^2 + \sum_{n=n_K}^N\vart1_n\Func0_0(\tilde{\varf2}_K) 
		\\+ \sum_{k=1}^K\sum_{n=n_{k-1}}^{n_k-1}\vart1_n\Func0_0(\tilde{\varf2}_K) 
		+ \IP{\varf1_{n_k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}.
	\end{multline}
	By \Cref{app: tn upper bound}, $\vart1_n\leq n$ so we can further simplify
	$$\sum_{n=\vars3}^{\vars4-1}\vart1_n \leq \sum_{n=\vars3}^{\vars4-1}n = (\vars4-\vars3)\frac{\vars4-1+\vars3}{2} \leq \frac{\vars4^2-\vars3^2}{2}$$
	to get the required bound.
\end{proof}

\subsection{Proof for Step 4}
\begin{lemma}[\Cref{thm: sufficiently fast}]\label{app:thm: sufficiently fast}
	\Paste{thm: sufficiently fast}
\end{lemma}
\todo{This proof is quite long and not very explicit. Should I compute the constant?}
\begin{proof}
	Inserting the assumed rates into \Cref{app:thm: mini FISTA convergence} gives
	\begin{equation}
		\vart1_N^2\Func0_0(\varf0_N) + \tfrac12\norm{\varf1_N-\tilde{\varf2}_K}^2 \lesssim \vars3_{\Varf0}^{2K} + (N+1)^2\vars3_{\Func0}^{-K} + \sum_{k=1}^{K}n_k^2\vars3_{\Func0}^{-k} + \vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}} + \vars3_{\Varf0}^{2k}.
	\end{equation}
	Each case now needs its own induction. When $\vars3_{\Varf0}>1$ we simplify the inequality to 
	\begin{align}
		\vart1_N^2\Func0_0(\varf0_N) + \tfrac12\norm{\varf1_N-\tilde{\varf2}_K}^2 &\lesssim \vars3_{\Varf0}^{2K} + \vars3_{\Varf0}^{2K+2} + \sum_{k=1}^{K}\vars3_{\Varf0}^{2k} + \vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}
		\\&\leq C_1\left(\vars3_{\Varf0}^{2K+2} + \frac{\vars3_{\Varf0}^{2K+2}}{\vars3_{\Varf0}^2-1} + \sum_{k=1}^{K}\vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}\right).
	\end{align}
	for some $C_1$ sufficiently large. Assume $\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}\leq C_2\vars3_{\Varf0}^k$ for $k\leq K$, then for $N=n_{K+1}-1$ we have
	\begin{equation}
		\tfrac12\norm{\varf1_{n_{K+1}-1}-\tilde{\varf2}_K}^2 \leq \frac{C_1\vars3_{\Varf0}^{2K+2}}{\vars3_{\Varf0}^2-1}\left(\vars3_{\Varf0}^2+C_2\right).
	\end{equation}
	If $C_2$ is sufficiently large, then $\frac{C_1\vars3_{\Varf0}^{2K+2}}{\vars3_{\Varf0}^2-1}\left(\vars3_{\Varf0}^2+C_2\right) \leq \tfrac12C_2^2\vars3_{\Varf0}^{2K+2}$ which completes the induction. This bounds the growth of the right hand side and so for any $N<n_{K+1}$ we have $t_N^2\Func0_0(\varf0_N) \lesssim \vars3_{\Varf0}^{2K}$.

>>>>

\paragraph{Step 5: Sufficiently slow refinement}
\begin{adaptiveStep}\label{step: step 5 sufficiently slow}
	The result of \Cref{step: step 4 sufficiently fast} is sufficient to prove convergence, but not yet a rate. If the subspaces refine too quickly, then the influence of $\norm{\varf0^*}=\infty$ will slow the rate of convergence. If the resolution is too coarse then we are overfitting to the discrete minimiser, but if the resolution is too fine then convergence of the energy is slow. \Cref{thm: sufficiently slow} balances these two factors in an optimal way for \cref{alg: refining FISTA} resulting in a convergence rate of 
	$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} \lesssim \frac{N^{2\kappa}}{N^2}$$
	for all $N\in\F N$, some $\kappa\in[0,1)$ depending on $\vars3_{\Varf0},\vars3_{\Func0}$. In particular, if $\varf0^*\in\F H$ then we recover the classical rate with $\kappa=0$.
\end{adaptiveStep}

@@

	Adaptively choosing the time of refinement is more challenging due to the non-descent property of FISTA. The idea is that $\spcf0^n$ should be refined as soon as the value $\Func0(\varf0_n)$ is sufficiently close to $\Func0(\varf2_n)$. This is analysed in \Cref{thm: stronger exponential FISTA convergence} which shows convergence results of the form
	$$ \min_{n\leq N}\Func0_0(\varf0_n) \lesssim \frac{N^{2\kappa}}{N^2}$$
	for all $N\in\F N$, with the same $\kappa\in[0,1)$ from \Cref{step: step 5 sufficiently slow}. The penalty for accelerating the refinement time is a potential loss of stability in $\Func0(\varf0_n)$, however, the asymptotic rate is equivalent and this behaviour has not been seen in numerical experiments.
\end{adaptiveStep}

@@

\begin{lemma}\label{thm: mini exponential FISTA convergence}
	\Copy{thm: mini exponential FISTA convergence}{
		Let $\varf0_n$, $\varf1_n$ be generated by \Cref{alg: refining FISTA}, $(n_k\in\F N)_{k=0}^\infty$ be a monotone increasing sequence, and choose
		$$\tilde{\varf2}_k \in \spcf0^{n_k} \cap \spcf0^{n_k+1}\cap \ldots \cap \spcf0^{n_{k+1}-1}$$
		for each $k\in\F N$. If such a sequence exists, then for all $K\in\F N$, $n_K\leq N<n_{K+1}$ we have
		\begin{multline*}
			\vart1_N^2\Func0_0(\varf0_N) + \sum_{n=1}^{N-1}\rho_n \Func0_0(\varf0_n) + \frac{\norm{\varf1_N-\tilde{\varf2}_K}^2}{2} \leq C + \frac{\norm{\tilde{\varf2}_K}^2}{2} + \frac{(N+1)^2-n_K^2}{2}\Func0_0(\tilde{\varf2}_K) \\+ \sum_{k=1}^K \frac{n_k^2-n_{k-1}^2}{2}\Func0_0(\tilde{\varf2}_{k-1}) + \IP{\varf1_{n_k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}
		\end{multline*}
		where $C = \frac{\norm{\varf0_0-\tilde{\varf2}_0}^2-\norm{\tilde{\varf2}_0}^2}{2}$.
	}
\end{lemma}
\noindent The proof is given in \cref{app:thm: mini exponential FISTA convergence}. The introduction of $n_k$ has greatly compressed the expression of \Cref{thm: mini FISTA convergence}. On the right-hand side, we now only consider $\Func0_0$ evaluated on the sequence $\tilde{\varf2}_k$ and there are $K$ elements to the sum rather than $N$.

@@

\begin{lemma}\label{thm: sufficiently fast}
	\Copy{thm: sufficiently fast}{
		Suppose $\spcf0^n, \varf0_n, \varf1_n$ and $n_k$ satisfy the conditions of \Cref{thm: mini exponential FISTA convergence} and $\{\tilde{\spcf0}^k\}$ forms an $(\vars3_{\Varf0},\vars3_{\Func0})$-discretisation for $\Func0$ with
		$$\tilde{\varf2}_k\in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k}\cap\spcf0^{n_k+1}\cap\ldots\cap\spcf0^{n_{k+1}-1}.$$
		If either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty,$
		\end{itemize}
		then
		$$ \Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2}$$
		for all $n_K\leq N< n_{K+1}$.}
\end{lemma}

@@

\begin{lemma}\label{thm: sufficiently slow}
	\Copy{thm: sufficiently slow}{
		Suppose $\varf0_n$ and $n_k$ are sequences satisfying 
		$$\Func0_0(\varf0_N) \lesssim \frac{\vars3_{\Varf0}^{2K}}{N^2} \qquad\text{where} \qquad  n_K^2\gtrsim \vars3_{\Func0}^K\vars3_{\Varf0}^{2K},$$
		then 
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}.$$}
\end{lemma}
\noindent The proof is given in \cref{app:thm: sufficiently slow}.

@@

		\bigbreak\noindent Suppose that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\simeq \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$,
		\end{itemize}
		then
		$$\Func0_0(\varf0_N) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}$$
		uniformly for $N\in\F N$.
	}
\end{theorem}
\noindent This theorem is very easy to implement and requires very little knowledge of how to estimate $\Func0_0(\varf0_n)$. So long as $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be computed analytically, choosing $\tilde{\spcf0}^k$ to be uniform discretisations and $\tilde{\spcf0}^k = \spcf0^{n_k}=\ldots= \spcf0^{n_{k+1}-1}$ will give the stated convergence rate.

@@

\Cref{thm: sufficiently fast} gives a sufficient condition on $n_k$ for converging at the rate $O(N^{2(\kappa-1)})$ but it is not necessary. Indeed for $n\leq n_k$ the result shows 
$$\Func0_0(\varf0_n) \geq \min_{\varf0\in\spcf0^n}\Func0_0(\varf0) = O(\vars3_{\Func0}^{-k}) = O(n^{2(\kappa-1)})$$
which suggests that to converge faster than $n^{2(\kappa-1)}$ requires choosing smaller $n_k$. As an example, in \Cref{sec: wavelet examples} we will see \cref{alg: refining FISTA} can converge at a linear rate, although this is not possible without adaptive refinement times. On the other hand, choice of spatial adaptivity has no impact on rate but can impact computational efficiency. It will be permitted to use greedy discretisation techniques so long as each collection $\{\spcf0^n \st n_k\leq n<n_{k+1}\}$ satisfies a monotonicity condition.

@@

		\bigbreak\noindent Suppose there exist $\vars3_{\Varf0},\vars3_{\Func0}\geq1$ such that either:
		\begin{itemize}
			\item $\vars3_{\Varf0}>1$ and $n_k^2\lesssim \vars3_{\Func0}^k\vars3_{\Varf0}^{2k}$,
			\item or $\vars3_{\Varf0}=1$, $\sum_{k=1}^\infty n_k^2\vars3_{\Func0}^{-k}<\infty$ and $\sum_{k=1}^\infty\norm{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}<\infty$
		\end{itemize}
		and both
		$$ \norm{\tilde{\varf2}_k}\lesssim \vars3_{\Varf0}^k \qquad\text{and}\qquad \Func0_0(\varf0_{n_K-1}) \lesssim \vars3_{\Func0}^{-K}.$$
		Whenever these conditions on $n_k$ are satisfied, then
		$$\min_{n\leq N}\Func0_0(\varf0_n) \lesssim \frac{1}{N^{2(1-\kappa)}}\qquad \text{ where } \qquad \kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2}$$
		uniformly for $N\in\F N$.
	}
\end{theorem}
\noindent The proof is given in \cref{app:thm: stronger exponential FISTA convergence}. If we directly compare \Cref{thm: exponential FISTA convergence,thm: stronger exponential FISTA convergence}, we note that the convergence rate is the same in both theorems but the price for better adaptivity is a slightly weaker stability guarantee (i.e. the addition of the min on the left-hand side). In \Cref{thm: exponential FISTA convergence}, as in the original FISTA algorithm, the sequence $\Func0(\varf0_n)$ is not monotone but the magnitude of oscillation is guaranteed to decay in time. This behaviour is boundedness is lost in \cref{thm: stronger exponential FISTA convergence} although it still holds on a subsequence. Although we do not prove it here, it can be shown that the stronger condition 
\begin{equation}
	\tilde{\varf2}_k \in \left[\argmin_{\varf0\in\tilde{\spcf0}^k}\Func0(\varf0)\right]\cap\spcf0^{n_k+1} \cap\ldots\cap \spcf0^{n_{k+1}-1}\cap\ldots\cap \spcf0^N
\end{equation}
is sufficient to restore the stronger guarantee on $\Func0_0(\varf0_n)$. Again, monotonicity of $\spcf0^n$ corresponds with improved stability of \cref{alg: refining FISTA}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General examples}\label{sec: general examples}
In this work we consider the Lasso problem to be both a motivating example and source of numerical examples, however, \Cref{alg: refining FISTA} is much more broadly applicable. The aim of this section is to justify this claim by demonstrating that the constants $\vars3_{\Varf0}$ and $\vars3_{\Func0}$ can be easily computed in many examples.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Inserting the computed rates into \Cref{thm: exponential FISTA convergence} or \Cref{thm: stronger exponential FISTA convergence} gives the guaranteed convergence rate
\begin{equation}\label{eq: Lasso energy rate}
	\kappa = \frac{\log \vars3_{\Varf0}^2}{\log \vars3_{\Func0} + \log \vars3_{\Varf0}^2} = \frac{d}{1+d} \quad \implies \quad \Func0_0(\varf0_n) \lesssim n^{-2(1-\kappa)} = n^{-\frac{2}{1+d}}.
\end{equation}
This rate can be used to infer the required resolution at each iteration, in particular on iteration $n$ with $n^2\sim (\vars3_{\Func0}\vars3_{\Varf0}^2)^k$ we expect the resolution to be
\begin{equation}\label{eq: Lasso resolution rate}
	2^{-k} = \left(\vars3_{\Func0}\vars3_{\Varf0}^2\right)^{\frac{k}{1+d}} \sim n^{-\frac{2}{1+d}}.
\end{equation}

@@

\begin{lemma}[\Cref{thm: mini exponential FISTA convergence}]\label{app:thm: mini exponential FISTA convergence}
	\Paste{thm: mini exponential FISTA convergence}
\end{lemma}
\begin{proof}
	This is just a telescoping of the right-hand side of \eqref{app:eq: FISTA inequality} with the introduction of $n_k$ and simplification $\varf2_n = \tilde{\varf2}_k$,
	\begin{multline}
		\tfrac12\norm{\varf2_N}^2 + \sum^N_{n=1} \vart1_n\Func0_0(\Pi_n\varf2_n) + \IP{\varf1_{n-1}}{\varf2_{n-1}-\varf2_n} = \tfrac12\norm{\tilde{\varf2}_K}^2 + \sum_{n=n_K}^N\vart1_n\Func0_0(\tilde{\varf2}_K) 
		\\+ \sum_{k=1}^K\sum_{n=n_{k-1}}^{n_k-1}\vart1_n\Func0_0(\tilde{\varf2}_{k-1}) 
		+ \IP{\varf1_{n_k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}.
	\end{multline}
	By \Cref{app: tn upper bound}, $\vart1_n\leq n$ so we can further simplify
	$$\sum_{n=\vars3}^{\vars4-1}\vart1_n \leq \sum_{n=\vars3}^{\vars4-1}n = (\vars4-\vars3)\frac{\vars4-1+\vars3}{2} \leq \frac{\vars4^2-\vars3^2}{2}$$
	to get the required bound.
\end{proof}

\subsection{Proof for Step 4}
\begin{lemma}[\Cref{thm: sufficiently fast}]\label{app:thm: sufficiently fast}
	\Paste{thm: sufficiently fast}
\end{lemma}
\begin{proof}
	Starting from \Cref{app:thm: mini exponential FISTA convergence} we get
	\begin{align}
		\vart1_N^2\Func0_0(\varf0_N) + \tfrac12\norm{\varf1_N-\tilde{\varf2}_K}^2 &\leq C + \frac{\norm{\tilde{\varf2}_K}^2}{2} + \frac{(N+1)^2-n_K^2}{2}\Func0_0(\tilde{\varf2}_K) + \sum_{k=1}^K \frac{n_k^2-n_{k-1}^2}{2}\Func0_0(\tilde{\varf2}_{k-1}) + \IP{\varf1_{n_k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}
		\\& \leq C + \frac{\norm{\tilde{\varf2}_K}^2}{2} + \frac{(N+1)^2}{2}\Func0_0(\tilde{\varf2}_K) \notag
		\\&\hspace{100pt}+\sum_{k=1}^K \frac{n_k^2}{2}\Func0_0(\tilde{\varf2}_{k-1}) + \IP{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}+\tilde{\varf2}_{k-1}}{\tilde{\varf2}_k-\tilde{\varf2}_{k+1}}.
	\end{align}
	The inductive step now depends on the value of $\vars3_{\Varf0}$.
	\begin{description}
		\item[Case $\vars3_{\Varf0}>1$:] We simplify the inequality
		\begin{align}
			\vart1_N^2\Func0_0(\varf0_N) + \tfrac12\norm{\varf1_N-\tilde{\varf2}_K}^2 &\lesssim \vars3_{\Varf0}^{2K} + (N+1)^2\vars3_{\Func0}^{-K} + \sum_{k=1}^{K}n_k^2\vars3_{\Func0}^{-k} + \vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}} + \vars3_{\Varf0}^{2k}
			\\&\leq C_1\left[\vars3_{\Varf0}^{2K+2} + \sum_{k=1}^{K}\vars3_{\Varf0}^{2k} + \vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}\right]
		\end{align}
		for some $C_1>C$. Choose $C_2\geq \norm{\varf1_{n_1-1}-\tilde{\varf2}_{1-1}}\vars3_{\Varf0}^{-1}$ such that 
		\begin{equation}
			\frac12C_2^2 \geq \frac{C_1}{\vars3_{\Varf0}^2-1}(C_2+\vars3_{\Varf0}^2).
		\end{equation}
		Assume $\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}\leq C_2\vars3_{\Varf0}^k$ for $1\leq k\leq K$ (trivially true for $K=1$), then for $N=n_{K+1}-1$ we have 
		\begin{align}
			\tfrac12\norm{\varf1_{n_{K+1}-1}-\tilde{\varf2}_K}^2 &\leq C_1\left[\vars3_{\Varf0}^{2K+2} + \sum_{k=1}^{K}\vars3_{\Varf0}^{2k} + \vars3_{\Varf0}^{k}\norm{\varf1_{n_k-1}-\tilde{\varf2}_{k-1}}\right]
			\\&\leq C_1\left[\vars3_{\Varf0}^{2K+2} + (1+C_2)\frac{\vars3_{\Varf0}^{2K+2}}{\vars3_{\Varf0}^2-1}\right]
			\\&\leq \frac{C_1\vars3_{\Varf0}^{2K+2}}{\vars3_{\Varf0}^2-1}\left(\vars3_{\Varf0}^2+C_2\right)
			\\&\leq \tfrac12(C_2\vars3_{\Varf0}^{K+1})^2.
		\end{align}
